---
title: "NHL Statistics"
output: 
  flexdashboard::flex_dashboard:
    theme: 
      version: 4
      bootswatch: yeti
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(polite)
library(purrr)
library(httr2)
library(httr)
library(leaflet)
library(shiny)
library(scales)
library(ggiraph)
```

About
=====================================  

    
Column {data-width=600}
-------------------------------------

Given the increased data availability and growing popularity of the National Hockey League (NHL), this project analyzes the league’s 32 hockey teams, with data dating back to 2022. With these statistics, we hope to answer key questions:

- How have teams’ success, as measured through winning percentage, goals, and points, shifted through the years?

- Which conferences are the most challenging to compete in?

- How much of a team’s success can be attributed to luck?

Our data comes from [Hockey Reference](https://www.hockey-reference.com/leagues/NHL_2026.html), a website that compiles hockey statistics for teams and players. In this project, we will focus on different ways of viewing a team’s success throughout time.

Source code for this dashboard can be found [here](https://github.com/ghoover24/Mini_Project_2).

Column {data-width=400}
-------------------------------------
(Image goes here)


```{r setup nhl data set, include=FALSE}
robotstxt::paths_allowed("https://www.hockey-reference.com/leagues/NHL_2026.html")

# Step One: scraping hockey website
nhl <- read_html("https://www.hockey-reference.com/leagues/NHL_2026.html")
str(nhl)

# Step Two: scraping tables from hockey website
table <- html_nodes(nhl, css = "table")
untidy_tables <- html_table(table, header = TRUE, fill = TRUE)

# function fro scraping multiple seasons
scrape <- function(x) {
  url <- read_html(str_c("https://www.hockey-reference.com/leagues/NHL_", x, ".html")) 
  table <- html_nodes(url, css = "table")
  untidy_tables <- html_table(table, header = TRUE, fill = TRUE) 
  eastern <- untidy_tables[[1]] |>                   # tidy eastern conference
    rename(Team = 1) |>                              # rename first column      
    slice(-1, -10)                                   # remove unnecessary rows
  western <- untidy_tables[[2]] |>                   # tidy western conference
    rename(Team = 1) |>                              # rename first column
    slice(-1, -10)                                   # remove unnecessary rows
  full_standings <- western |>                       # join eastern and western
    full_join(eastern) |>
    mutate(season = x)
  full_standings
}

nhl_full_years <- map_df(2022:2026, scrape)

# clean names and levels
nhl_full_years <- nhl_full_years |>
  rename(PTS_PCT = `PTS%`,             # make tidy
         RPt_PCT = `RPt%`,
         RgPT_PCT = `RgPt%`) |>
  mutate(across(GP:RW, as.numeric),
         RgPT_PCT = as.numeric(RgPT_PCT),
         Team = str_replace(Team, "\\*", ""),
         W_PCT = (W / GP),          # create new vars
         W_PCT = round(W_PCT, 4))

# Create df to sort conference and divisions
nhl_conferences <- data.frame(
  Team = c("Boston Bruins", "Buffalo Sabres", "Detroit Red Wings", "Florida Panthers", "Montreal Canadiens", "Ottawa Senators", "Tampa Bay Lightning", "Toronto Maple Leafs","Carolina Hurricanes", "Columbus Blue Jackets", "New Jersey Devils", "New York Islanders", "New York Rangers", "Philadelphia Flyers", "Pittsburgh Penguins", "Washington Capitals", "Chicago Blackhawks", "Colorado Avalanche", "Dallas Stars", "Minnesota Wild", "Nashville Predators", "St. Louis Blues", "Utah Mammoth", "Winnipeg Jets", "Anaheim Ducks", "Calgary Flames", "Edmonton Oilers", "Los Angeles Kings", "San Jose Sharks", "Seattle Kraken", "Vancouver Canucks", "Vegas Golden Knights"),
  conference = c(rep("Eastern", 16), rep("Western", 16)),
  division = c(rep("Atlantic", 8), rep("Metropolitan", 8), rep("Central", 8), rep("Pacific", 8)))


# join dfs to include all necessary variables
nhl_all <- nhl_full_years |>
  left_join(nhl_conferences) # add conferences and divisions
```

```{r location data, include=FALSE}
robotstxt::paths_allowed("https://en.wikipedia.org/wiki/List_of_National_Hockey_League_arenas")

# Scraping arenas and team name
locations <- read_html("https://en.wikipedia.org/wiki/List_of_National_Hockey_League_arenas")
str(locations)

table <- html_nodes(locations, css = "table")                   # scrape tables from wiki
untidy_tables <- html_table(table, header = TRUE, fill = TRUE)  

location_tbl <- untidy_tables[[1]] |>           # 
  rename(Image = 1, arena = 2, location = 3,
         team = 4, capacity = 5, opened = 6,
         season_one = 7, refs = 8) |>
  select(arena, team) |>
  mutate(arena = str_replace(arena, " ", "_"),
         arena = str_replace(arena, " ", "_"))

# test scrape coordinates
coord <- read_html("https://en.wikipedia.org/wiki/PPG_Paints_Arena")
lat_temp <- html_nodes(coord, ".rt-overlay , #mw-content-text .latitude")
lat_text <- html_text(lat_temp)

long_temp <- html_nodes(coord, ".rt-overlay , #mw-content-text .longitude")
long_text <- html_text(long_temp)

coordinates <- tibble(lat = lat_text,
                      long = long_text)

# create arena list
arenas_col <- location_tbl |> 
  select(arena) |>
  mutate(arena = str_replace(arena, " ", "_"),
         arena = str_replace(arena, " ", "_"))
arena <-as.list(arenas_col)

# for loop for scraping coordinates
coordinates <- tibble(latitude = character(), longitude = character(), arena = character())

for(i in 1:32) {
  url <- str_c("https://en.wikipedia.org/wiki/", arena[[1]][i])
  one_arena <- GET(url)
  coord <- read_html(one_arena)
# latitude scrape
  lat_temp <- html_nodes(coord, ".rt-overlay , #mw-content-text .latitude")
  lat_text <- html_text(lat_temp)
# longitude scrape
  long_temp <- html_nodes(coord, ".rt-overlay , #mw-content-text .longitude")
  long_text <- html_text(long_temp)

  coordinates[i,1] <- lat_text[1]
  coordinates[i,2] <- long_text[1]
  coordinates[i,3] <- arena[[1]][i]
}

# join two dfs
locations <- location_tbl |>
  full_join(coordinates) |>
  mutate(arena = str_replace(arena, "_", " "), # clean arena names
         arena = str_replace(arena, "_", " "))

# clean latitude and longitude for map
fix_lat_long <- function(x) {
  degree  <- as.numeric(str_extract(x, "^[0-9.]+"))     # remove non digits
  minutes <- as.numeric(str_extract(x, "[0-9.]+(?=′)")) 
  seconds <- as.numeric(str_extract(x, "[0-9.]+(?=″)"))
  decimal <- degree + minutes/60 + seconds/3600
  return(decimal)
}

locations_clean <- locations |>
  mutate(latitude = sapply(latitude, fix_lat_long),
    longitude = sapply(longitude, fix_lat_long),
    longitude = as.numeric(str_c("-",longitude)))

# prepare locations data for joining to NHl data
locations_clean <- locations_clean |> 
  rename(Team = team)                 # match variable name

# create specific df for leaflet map
nhl_map <- nhl_all |>
  full_join(locations_clean) |>
    mutate(W_PCT = W_PCT * 100,
         W_PCT_map_size = (as.numeric(W) / as.numeric(GP)) * 30,
         popup = str_c(arena, ": the ", Team, " have a win percentage of ", W_PCT, "%")) 
```


Map of Teams
=====================================

Column
-------------------------------------

```{r, echo = FALSE}
nhl_map_26 <- nhl_map |>
  filter(season == 2026)

bins <- c(0, 30, 40, 50, 60, 70, 80, 100)                      # for color/fill
pal <- colorBin("YlOrRd", domain = nhl_map$W_PCT, bins = bins) # for color/fill

leaflet() |>
  setView(-97, 40, 4) |>
  addTiles() |>
  addCircleMarkers(data = nhl_map_26,  # circle markers for each team/arena
        lat = ~ latitude, 
        lng = ~ longitude,
        popup = ~ popup,
        radius = ~ W_PCT_map_size,     # larger circles for teams with higher win percentages
        weight = 4,
        color = "grey",
        fillColor = ~pal(W_PCT),       # darker red for teams with higher win percentages
        fillOpacity = 1)
```

Column
-------------------------------------

From the map you can see geographically where each teams’ home stadium resides. There is a general trend for many teams to be located near coasts with many gaps within the Midwest and Pacific Northwest. Most of the arenas are clustered in the Northeast region in small, densely populated states with larger hockey cultures.

The gaps seen in the middle of the U.S. could be attributed to the smaller populations and lack of large cities in these states. These factors possibly limit the amount of fans who can reasonably attend home games, decreasing the profits that can be made off the team. This lower level of profit is most likely what sways the NHL to not invest in placing teams in these regions.

It is also interesting to note that the Canadian teams are fairly close to the American border compared to the American teams. This trend could also be attributed to the amount of profits that can be made off of these teams. These teams are located in the most populated cities in Canada as most people living in Canada live in the southern region of the country. This leaves more people available to attend games at home for these teams. Being closer to the border is also beneficial as it leaves Americans in the northern U.S. able to attend away games without travelling too far. 



Team Success {data-orientation=rows}
=====================================     

Row
-----------------------------------------------------------------------

```{r, echo=FALSE}
inputPanel(
  selectInput("chosen_stat", label = "Choose a Statistic:", # users can choose a statistic for line graph
              choices = c("Number of Wins" = "W",           # nicer names for selection
                          "Number of Losses" = "L",
                          "Win Percentage" = "W_PCT",
                          "Points" = "PTS",
                          "Goals For" = "GF",
                          "Goals Against" = "GA")),
  selectInput("chosen_team", label = "Choose Teams:",       # users can choose teams to compare
              choices = c(nhl_all$Team), 
              selected = c("Minnesota Wild", "Colorado Avalanche", "Dallas Stars"), multiple = TRUE) # allow multiple to be selected
)

renderPlot(
  nhl_all |>
    filter(Team %in% input$chosen_team) |> # filter teams to the selected teams
    ggplot(aes(x = season, y = .data[[input$chosen_stat]], color = Team)) + # y-axis is chosen statistic
    geom_point() + # each point represents a team at a specific season
    geom_smooth() + 
    labs(x = "Season",
         y = "Chosen Statistic") +
    theme_classic() # nice theme
)
```

Row
-----------------------------------------------------------------------

Over the course of many seasons, each teams’ successes can drastically change due to many different factors. This plot allows you to see these trends over the past five seasons and compare them across teams. For example, in the default settings, you can see how the total number of games three teams have won since the 2021-2022 season. In the previous three seasons, the Colorado Avalanche’s wins have remained relatively steady and increased by five games for the most recent season (2025-2026). The Dallas Stars have also had a stable number of wins across seasons. The Minnesota Wild is seen to have a fluctuating number of wins with a dramatic decrease from the 2021-2022 season to the 2023-2024 season. However, the Wild are on a general incline in wins. 


Conference Strength
=====================================   

Row
-------------------------------------

```{r density plot, echo=FALSE}
inputPanel(
  selectInput("year", label = "Choose a Season:",
              choices = c(2022, 2023, 2024, 2025, 2026))     #allows user to select a season
)

renderPlot(
  nhl_all |>
    rename(Division = division) |>         #fixes the title for the legend
    filter(!is.na(conference),             #eliminates teams that no longer exists
           season == input$year) |>
    ggplot(aes(x = W_PCT, color = Division, fill = Division)) +
    geom_density(bw = "nrd0", alpha = 0.4) +
    labs(x = "Winnning Percentage", y = "Density") +
    theme_classic()
)
```

In the NHL, intradivisional games are frequent between teams. Due to differences in teams’ skill levels within the divisions, a certain division may be more difficult than another. These variations in ability create a reality that teams of a certain division may experience greater competitive challenges than the team of another division. Sports statisticians describe this phenomenon as “strength of schedule.”

This density plot shows the differences between the divisions through the past five NHL seasons. Generally speaking, the central division experienced a lot of success from 2022 through 2024, which means that the average team would be less likely to succeed in that division during that time frame. However, the central division appears to have tailed off slightly in 2026, with a distribution of winning percentages centered below .500.

Column
-------------------------------------

### Atlantic Division
```{r, echo = FALSE}
Atlantic_table <- data.frame(nhl_all) |>
  filter(!is.na(conference)) |>
  filter(division == "Atlantic") |>
  group_by(Team) |>
  summarize(W = sum(W),
            L = sum(L),
            GP = sum(GP),
            Win_Pct = round((W/GP), 3)) |>
  arrange(desc(Win_Pct))
knitr::kable(Atlantic_table)
```
### Central Division
```{r, echo = FALSE}
Central_table <- data.frame(nhl_all) |>
  filter(!is.na(conference)) |>
  filter(division == "Central") |>
  group_by(Team) |>
  summarize(W = sum(W),
            L = sum(L),
            GP = sum(GP),
            Win_Pct = round((W/GP), 3)) |>
  arrange(desc(Win_Pct))
knitr::kable(Central_table)
```
### Metropolitan Division
```{r, echo = FALSE}
Metropolitan_table <- data.frame(nhl_all) |>
  filter(!is.na(conference)) |>
  filter(division == "Metropolitan") |>
  group_by(Team) |>
  summarize(W = sum(W),
            L = sum(L),
            GP = sum(GP),
            Win_Pct = round((W/GP), 3)) |>
  arrange(desc(Win_Pct))
knitr::kable(Metropolitan_table)
```
### Pacific Division
```{r, echo = FALSE}
Pacific_table <- data.frame(nhl_all) |>
  filter(!is.na(conference)) |>
  filter(division == "Pacific") |>
  group_by(Team) |>
  summarize(W = sum(W),
            L = sum(L),
            GP = sum(GP),
            Win_Pct = round((W/GP), 3)) |>
  arrange(desc(Win_Pct))
knitr::kable(Pacific_table)
```

Team Luck
=====================================   

Column
-------------------------------------

```{r, echo = FALSE}
inputPanel(
  selectInput("team_select", label = "Choose a team:",
              choices = c(nhl_all$Team), selected = c("Minnesota Wild")
))

renderPlot(
nhl_all |> 
    mutate(selected = Team == input$team_select) |>
    ggplot(aes(x = W_PCT, y = SRS)) + 
    geom_point(aes(color = selected, size = selected), show.legend = FALSE) +
    geom_smooth(method = lm, se = FALSE) +
    labs(y = "Simple Rating System (SRS)", x = "Winning Percentage") +
    theme_classic()
)
```

Column
-------------------------------------

The Simple Rating System (SRS) statistic attempts to quantify the true skill level of a team, determined through their strength of schedule, goals scored, and goals against. In theory, teams that score many goals and allow few goals should win a lot of games. However, the strength of a team’s opponent also impacts their ability to succeed. Therefore, teams that win games while playing more challenging opponents will get a boost to their SRS score. In theory, this statistic should be a simple way to show us the true skill level of a team.

In this scatterplot, we ran a linear regression between SRS and winning percentage. With an R^2 value of .92, the regression found a very strong relationship between the variables. The linear regression line between the variables shows the expected winning percentage, given a team’s SRS score. Teams that win more than expected, based on their SRS score could be considered ‘lucky,’ according to the metric, while those who won fewer games than expected could be considered ‘unlucky.’

Insights into the true skill of a team is valuable because it allows front offices and fans to create stronger expectations for their future performance. If a team is winning more games than expected, fans and analysts may set their expectations lower to match what the team is expected to do. For example, through each of the past five seasons, the Minnesota Wild has sat either at or below expectations, meaning that their fans could expect their team to win slightly more games than they have been winning, assuming the team maintains its roster, continues to perform with similar success, and plays against a league-average team.

### Conclusion

Each team in the NHL has had a unique story from the past five years. While some have improved, others have declined. Much of this can be attributed to their individual success as a team, but team performance also depends on the competitive environment they play in. Taking divisional success into account provides a clearer picture of the true success of a team. When fans and analysts evaluate teams, it is important to recognize a team’s trends in success, and the competitive atmosphere they exist within.

